# Post-Validation Filtering {#filtering}

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
dat <- read_csv("~/Desktop/master_output_data_analysis_v15.csv")
```

## Fixing algorithm errors discovered post-validation
Here we hard-coded several specific fixes to the classification that repeatedly cropped up during our validation process. All have been checked to have 0 mis-classifications.
```{r}
# Removing grants containing "dinner"
dat <- dat %>%
  filter(str_detect(Description, "dinner") == FALSE)

# Classifying "dissemination" as act_edu
sum(dat$cat_act_edu)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "dissemination") == TRUE) {
    dat$cat_act_edu[i] = 1
  }
}
sum(dat$cat_act_edu)

# Removing threat_energytransport classification for "palm oil"
sum(dat$cat_threat_energytransport)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "palm oil|oil-palm|oil palm") == TRUE) {
    dat$cat_threat_energytransport[i] = 0
  }
}
sum(dat$cat_threat_energytransport)

# Removing hab_other classification for "aquatic" (will be freshwater/marine)
sum(dat$cat_habitat_other)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "aquatic") == TRUE) {
    dat$cat_habitat_other[i] = 0
  }
}
sum(dat$cat_habitat_other)

```

## Remove grants with no associated dollar amount
Those grants with no funding amount recorded contributed nothing to this analysis, and were removed.
```{r}
dat.1 <- dat %>% 
  drop_na(Grant.Amount) %>%
  dplyr::select(1:42) 
names(dat.1)
```

## Filtering Aquarium/Zoo Grants

Any funding from grants classified as Aquarium/Zoo, regardless of other categories it was also associated with, was limited to to the Aquarium/Zoo category. **This is because...**

There was one exception to this: if classified into social justice, the the funding was counted in that category. **This is because...**

```{r}
for(i in 1:nrow(dat.1)){
  if(dat.1$cat_aquarium_zoo[i]==1){
    dat.1[i,c(11:17, 19:28, 30:42)]=0
  }
}
```
