# Multi-category Filtering {#multiCatFilter}

The code below governs the way we dealt with multiple, potentially overlapping categories on individual grants. The general format is as follows:  

1. Grants with >2 independent classfications (i.e. habitat_terrestrial, habitat_marine, and habitat_freshwater) are reclassified as "other" within their broader categories (i.e. habitat_other).   
2. Grants with 2 independent classifications (i.e. habitat_terrestrial and habitat_marine) are classified into a separate, joint classification within their broader categories (i.e. habitat_MarineTerrestrial).  
3. Those grants with shared independent and joint classifications default to the joint category described above.  

## Adding Category Sum columns

The above rules are built on dummy variables that take the sum of our binary classification data across the columns of each broader category.
```{r}
dat.1 <- dat.1 %>%
  mutate(Num.Class = rowSums(.[,11:length(.)])) %>%
  mutate(Num.Thr = rowSums(.[37:42])) %>%
  mutate(Num.Act = rowSums(.[11:17])) %>%
  mutate(Num.Hab = rowSums(.[c(25,26,28)])) %>%
  mutate(Num.Tax = rowSums(.[30:36])) %>%
  mutate(Num.Bio = rowSums(.[19:20]))
```

## Export Multi-category file

The foundation of this multi-category dataset is exported prior to the full multi-category processing to provide a dataset more suited to the network analysis.
```{r}
# Export file before category mutli-category processing for network analysis
dat.1.multi <- dat.1
write_csv(dat.1.multi, "~/Desktop/master_output_data_analysis_v15_mutlicategories.csv")
```

### Multi-category Filtering: Habitat Categories
```{r}
hab <- dat.1[,c(1:10,22:28)]

# Marine-Terrestrial
for(i in 1:nrow(hab)){
  if(hab$cat_habitat_marine[i]==1 & hab$cat_habitat_terrestrial[i]==1) {
    hab$cat_habitat_MarineTerrestrial[i]=1}
  if(hab$cat_habitat_MarineTerrestrial[i] == 1) {
    hab$cat_habitat_marine[i]=0
    hab$cat_habitat_terrestrial[i]=0
  }
}

# Marine-Freshwater
for(i in 1:nrow(hab)){
  if(hab$cat_habitat_marine[i]==1 & hab$cat_habitat_freshwater[i]==1) {
    hab$cat_habitat_FreshwaterMarine[i]=1}
  if(hab$cat_habitat_FreshwaterMarine[i] == 1) {
    hab$cat_habitat_marine[i]=0
    hab$cat_habitat_freshwater[i]=0
  }
}

# Freshwater-Terrestrial  
for(i in 1:nrow(hab)){
  if(hab$cat_habitat_terrestrial[i]==1 & hab$cat_habitat_freshwater[i]==1) {
    hab$cat_habitat_FreshwaterTerrestrial[i]=1}
  if(hab$cat_habitat_FreshwaterTerrestrial[i] == 1) {
    hab$cat_habitat_terrestrial[i]=0
    hab$cat_habitat_freshwater[i]=0
  }
}

## Check
sum(hab$cat_habitat_other)
sum(hab$cat_habitat_terrestrial)

# Habitat other (>2 categories)
for(i in 1:nrow(hab)){
  if(sum(hab[i,c(11:15,17)]) > 1){
    hab$cat_habitat_other[i] = 1
    hab[i, c(11:15,17)] = 0
  }
}

## Check
sum(hab$cat_habitat_other)
sum(hab$cat_habitat_terrestrial)


# If hab_other, no other habitat classification
for(i in 1:nrow(hab)){
  if(hab$cat_habitat_other[i]==1) {
    hab[i,c(11:15,17)] = 0
  }
}  

## Check
sum(hab$cat_habitat_other)
sum(hab$cat_habitat_terrestrial)
sum(dat.1$cat_habitat_freshwater)
sum(hab$cat_habitat_freshwater) # Huge change in freshwater number here, showing freshwater classified with other categories a lot


### Join filtered habitat columns with original data
data.final <- dat.1 
names(data.final)
data.final[,22:28] <- hab[,11:17]

# Check
sum(dat.1$cat_habitat_freshwater)
sum(data.final$cat_habitat_freshwater)
```

### Multi-category Filtering: Bio Categories
```{r}
data.final <- data.final %>%
  mutate(cat_bio_ecosys_species = 0)

## Filling cells

#test
sum(data.final$cat_bio_ecosys)

for(i in 1:nrow(data.final)){
  if(data.final$cat_bio_ecosys[i]==1 & data.final$cat_bio_species[i]==1) {
    data.final$cat_bio_ecosys_species[i]=1
    data.final$cat_bio_ecosys[i]=0
    data.final$cat_bio_species[i]=0
  }
}

#retest
sum(data.final$cat_bio_ecosys)
sum(data.final$cat_bio_ecosys_species)
```

### Multi-category Filtering: Action Categories
```{r}
act <- data.final[,c(1:17)]

act <- act %>% 
  mutate(sumAcross = cat_act_area + cat_act_community + cat_act_edu + cat_act_funding + 
           cat_act_policy + cat_act_research + cat_act_restor) %>%
  filter(sumAcross > 0)

# Checking number of grants/money for multiple action category grants
tot_act <- sum(act$Grant.Amount) # total = 

act5 <- act %>%
  filter(sumAcross <= 5)
nrow(act5)/nrow(act)
sum(act5$Grant.Amount)/tot_act
# Retained 99.95% of grants, 99.98% of money

act4 <- act %>%
  filter(sumAcross <= 4)
nrow(act4)/nrow(act)
sum(act4$Grant.Amount)/tot_act
# Retained 99.46% of grants, 99.70% of money

act3 <- act %>%
  filter(sumAcross <= 3)
nrow(act3)/nrow(act)
sum(act3$Grant.Amount)/tot_act
# Retained 97.63% of grants, 98.64% of money

act2 <- act %>%
  filter(sumAcross <= 2)
nrow(act2)/nrow(act)
sum(act2$Grant.Amount)/tot_act
# Retained 90.97% of grants, 91.96% of money

act1 <- act %>%
  filter(sumAcross <= 1)
nrow(act1)/nrow(act)
sum(act1$Grant.Amount)/tot_act
# Retained 67.42% of grants, 67.28% of money


n_actions <- 1:5
n_act_grants <- c(nrow(act1), 
              nrow(act2) - nrow(act1), 
              nrow(act3) - nrow(act2),
              nrow(act4) - nrow(act3),
              nrow(act5) - nrow(act4))

actions <- data.frame(n_actions, n_act_grants)
ggplot(actions, aes(x = n_actions, y = n_act_grants)) +
  geom_line() +
  labs(title = "# Grants in Multiple Action Categories")


## >2 category grants into (act_other)

act_other <- data.final %>% 
  mutate(sum_act = cat_act_area + cat_act_community + cat_act_edu + cat_act_funding + 
           cat_act_policy + cat_act_research + cat_act_restor) %>%
  filter(sum_act >2)


## Adding 2-action categories and other
data.final <- data.final %>%
  mutate(cat_act_area_comm = 0,
         cat_act_area_edu = 0,
         cat_act_area_funding = 0,
         cat_act_area_policy = 0,
         cat_act_area_research = 0,
         cat_act_area_restor = 0,
         cat_act_community_edu = 0,
         cat_act_community_funding = 0,
         cat_act_community_policy = 0,
         cat_act_community_research = 0,
         cat_act_community_restor = 0,
         cat_act_edu_funding = 0,
         cat_act_edu_policy = 0,
         cat_act_edu_research = 0,
         cat_act_edu_restor = 0,
         cat_act_funding_policy = 0,
         cat_act_funding_research = 0,
         cat_act_funding_restor = 0,
         cat_act_policy_research = 0,
         cat_act_policy_restor = 0,
         cat_act_research_restor = 0,
         cat_act_other = 0)


#### Filling Multiple Action cells ####

## Other Actions (>2 classifications)
#test
sum(data.final$cat_act_research)

for(i in 1:nrow(data.final)) {
  if(sum(data.final[i,11:17]) > 2) {
    data.final$cat_act_other[i] = 1
    data.final[i,11:17] = 0
  }
}

#retest
sum(data.final$cat_act_research)
sum(data.final$cat_act_other) # matches nrow of act_other DF


## Area + Community
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_community[i]==1) {
    data.final$cat_act_area_comm[i]=1}
  if(data.final$cat_act_area_comm[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_community[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_comm)


## Area + Education
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_edu[i]==1) {
    data.final$cat_act_area_edu[i]=1}
  if(data.final$cat_act_area_edu[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_edu[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_edu)


## Area + Funding
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_funding[i]==1) {
    data.final$cat_act_area_funding[i]=1}
  if(data.final$cat_act_area_funding[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_funding[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_funding)


## Area + Policy
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_policy[i]==1) {
    data.final$cat_act_area_policy[i]=1}
  if(data.final$cat_act_area_policy[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_policy[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_policy)


## Area + Research
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_research[i]==1) {
    data.final$cat_act_area_research[i]=1}
  if(data.final$cat_act_area_research[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_research[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_research)


## Area + Restoration
#test
sum(data.final$cat_act_area)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_area[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_area_restor[i]=1}
  if(data.final$cat_act_area_restor[i]==1) {
    data.final$cat_act_area[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_area)
sum(data.final$cat_act_area_restor)


## Community + Education
#test
sum(data.final$cat_act_community)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_community[i]==1 & data.final$cat_act_edu[i]==1) {
    data.final$cat_act_community_edu[i]=1}
  if(data.final$cat_act_community_edu[i]==1) {
    data.final$cat_act_community[i]=0
    data.final$cat_act_edu[i]=0
  }
}

#retest
sum(data.final$cat_act_community)
sum(data.final$cat_act_community_edu)


## Community + Funding
#test
sum(data.final$cat_act_community)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_community[i]==1 & data.final$cat_act_funding[i]==1) {
    data.final$cat_act_community_funding[i]=1}
  if(data.final$cat_act_community_funding[i]==1) {
    data.final$cat_act_community[i]=0
    data.final$cat_act_funding[i]=0
  }
}

#retest
sum(data.final$cat_act_community)
sum(data.final$cat_act_community_funding)


## Community + Policy
#test
sum(data.final$cat_act_community)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_community[i]==1 & data.final$cat_act_policy[i]==1) {
    data.final$cat_act_community_policy[i]=1}
  if(data.final$cat_act_community_policy[i]==1) {
    data.final$cat_act_community[i]=0
    data.final$cat_act_policy[i]=0
  }
}

#retest
sum(data.final$cat_act_community)
sum(data.final$cat_act_community_policy)


## Community + Research
#test
sum(data.final$cat_act_community)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_community[i]==1 & data.final$cat_act_research[i]==1) {
    data.final$cat_act_community_research[i]=1}
  if(data.final$cat_act_community_research[i]==1) {
    data.final$cat_act_community[i]=0
    data.final$cat_act_research[i]=0
  }
}

#retest
sum(data.final$cat_act_community)
sum(data.final$cat_act_community_research)


## Community + Restoration
#test
sum(data.final$cat_act_community)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_community[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_community_restor[i]=1}
  if(data.final$cat_act_community_restor[i]==1) {
    data.final$cat_act_community[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_community)
sum(data.final$cat_act_community_restor)


## Education + Funding
#test
sum(data.final$cat_act_edu)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_edu[i]==1 & data.final$cat_act_funding[i]==1) {
    data.final$cat_act_edu_funding[i]=1}
  if(data.final$cat_act_edu_funding[i]==1) {
    data.final$cat_act_edu[i]=0
    data.final$cat_act_funding[i]=0
  }
}

#retest
sum(data.final$cat_act_edu)
sum(data.final$cat_act_edu_funding)


## Education + Policy
#test
sum(data.final$cat_act_edu)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_edu[i]==1 & data.final$cat_act_policy[i]==1) {
    data.final$cat_act_edu_policy[i]=1}
  if(data.final$cat_act_edu_policy[i]==1) {
    data.final$cat_act_edu[i]=0
    data.final$cat_act_policy[i]=0
  }
}

#retest
sum(data.final$cat_act_edu)
sum(data.final$cat_act_edu_policy)


## Education + Research
#test
sum(data.final$cat_act_edu)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_edu[i]==1 & data.final$cat_act_research[i]==1) {
    data.final$cat_act_edu_research[i]=1}
  if(data.final$cat_act_edu_research[i]==1) {
    data.final$cat_act_edu[i]=0
    data.final$cat_act_research[i]=0
  }
}

#retest
sum(data.final$cat_act_edu)
sum(data.final$cat_act_edu_research)


## Education + Restoration
#test
sum(data.final$cat_act_edu)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_edu[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_edu_restor[i]=1}
  if(data.final$cat_act_edu_restor[i]==1) {
    data.final$cat_act_edu[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_edu)
sum(data.final$cat_act_edu_restor)


## Funding + Policy
#test
sum(data.final$cat_act_funding)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_funding[i]==1 & data.final$cat_act_policy[i]==1) {
    data.final$cat_act_funding_policy[i]=1}
  if(data.final$cat_act_funding_policy[i]==1) {
    data.final$cat_act_funding[i]=0
    data.final$cat_act_policy[i]=0
  }
}

#retest
sum(data.final$cat_act_funding)
sum(data.final$cat_act_funding_policy)


## Funding + Research
#test
sum(data.final$cat_act_funding)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_funding[i]==1 & data.final$cat_act_research[i]==1) {
    data.final$cat_act_funding_research[i]=1}
  if(data.final$cat_act_funding_research[i]==1) {
    data.final$cat_act_funding[i]=0
    data.final$cat_act_research[i]=0
  }
}

#retest
sum(data.final$cat_act_funding)
sum(data.final$cat_act_funding_research)


## Funding + Restoration
#test
sum(data.final$cat_act_funding)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_funding[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_funding_restor[i]=1}
  if(data.final$cat_act_funding_restor[i]==1) {
    data.final$cat_act_funding[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_funding)
sum(data.final$cat_act_funding_restor)


## Policy + Research
#test
sum(data.final$cat_act_policy)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_policy[i]==1 & data.final$cat_act_research[i]==1) {
    data.final$cat_act_policy_research[i]=1}
  if(data.final$cat_act_policy_research[i]==1) {
    data.final$cat_act_policy[i]=0
    data.final$cat_act_research[i]=0
  }
}

#retest
sum(data.final$cat_act_policy)
sum(data.final$cat_act_policy_research)


## Policy + Restoration
#test
sum(data.final$cat_act_policy)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_policy[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_policy_restor[i]=1}
  if(data.final$cat_act_policy_restor[i]==1) {
    data.final$cat_act_policy[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_policy)
sum(data.final$cat_act_policy_restor)


## Research + Restoration
#test
sum(data.final$cat_act_research)

for(i in 1:nrow(data.final)){
  if(data.final$cat_act_research[i]==1 & data.final$cat_act_restor[i]==1) {
    data.final$cat_act_research_restor[i]=1}
  if(data.final$cat_act_research_restor[i]==1) {
    data.final$cat_act_research[i]=0
    data.final$cat_act_restor[i]=0
  }
}

#retest
sum(data.final$cat_act_research)
sum(data.final$cat_act_research_restor)
```

### Multi-category Filtering: Threat Categories
```{r}
threat <- data.final[,c(1:10, 37:42)]

threat <- threat %>% 
  mutate(sumAcross = cat_threat_climate + cat_threat_energytransport + 
           cat_threat_habitat + cat_threat_inv_spec + cat_threat_overexp + 
           cat_threat_pollution) %>%
  filter(sumAcross > 0)

# Checking number of grants/money for multiple threation category grants
tot_threat <- sum(threat$Grant.Amount) # total = 

threat4 <- threat %>%
  filter(sumAcross <= 4)
nrow(threat4)/nrow(threat)
sum(threat4$Grant.Amount)/tot_threat
# Retained all grants

threat3 <- threat %>%
  filter(sumAcross <= 3)
nrow(threat3)/nrow(threat)
sum(threat3$Grant.Amount)/tot_threat
# Retained 99.75% of grants, 99.21% of money

threat2 <- threat %>%
  filter(sumAcross <= 2)
nrow(threat2)/nrow(threat)
sum(threat2$Grant.Amount)/tot_threat
# Retained 96.95% of grants, 95.39% of money

threat1 <- threat %>%
  filter(sumAcross <= 1)
nrow(threat1)/nrow(threat)
sum(threat1$Grant.Amount)/tot_threat
# Retained 77.55% of grants, 72.19% of money

n_threats <- 1:4
n_threat_grants <- c(nrow(threat1), 
                  nrow(threat2) - nrow(threat1), 
                  nrow(threat3) - nrow(threat2),
                  nrow(threat4) - nrow(threat3))

threats <- data.frame(n_threats, n_threat_grants)
ggplot(threats, aes(x = n_threats, y = n_threat_grants)) +
  geom_line() +
  labs(title = "# Grants in Multiple Threat Categories")

#### >2 category grants (threat_other)

threat_other <- data.final %>% 
  mutate(sum_threat = cat_threat_climate + cat_threat_energytransport + 
           cat_threat_habitat + cat_threat_inv_spec + cat_threat_overexp + 
           cat_threat_pollution) %>%
  filter(sum_threat >2)

#### Adding 2-threat categories

data.final <- data.final %>%
  mutate(cat_threat_climate_energytransport = 0,
         cat_threat_climate_habitat = 0,
         cat_threat_climate_inv_spec = 0,
         cat_threat_climate_overexp = 0,
         cat_threat_climate_pollution = 0,
         cat_threat_energytransport_habitat = 0,
         cat_threat_energytransport_inv_spec = 0,
         cat_threat_energytransport_overexp = 0,
         cat_threat_energytransport_pollution = 0,
         cat_threat_habitat_inv_spec = 0,
         cat_threat_habitat_overexp = 0,
         cat_threat_habitat_pollution = 0,
         cat_threat_inv_spec_overexp = 0,
         cat_threat_inv_spec_pollution = 0,
         cat_threat_overexp_pollution = 0,
         cat_threat_other = 0)



#### Filling Mulitple Threat cells ####

## Other Threats (>2 Categories)
#test
sum(data.final$cat_threat_overexp)

for(i in 1:nrow(data.final)){
  if(sum(data.final[i,37:42]) > 2) {
    data.final$cat_threat_other[i] = 1
    data.final[i, 37:42] = 0
  }
}

#retest
sum(data.final$cat_threat_overexp)
sum(data.final$cat_threat_other) # matches nrow of threats_other DF


## Climate + Energy/Transportation
#test
sum(data.final$cat_threat_climate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_climate[i]==1 & data.final$cat_threat_energytransport[i]==1) {
    data.final$cat_threat_climate_energytransport[i]=1}
  if(data.final$cat_threat_climate_energytransport[i]==1) {
    data.final$cat_threat_climate[i]=0
    data.final$cat_threat_energytransport[i]=0
  }
}

#retest
sum(data.final$cat_threat_climate)
sum(data.final$cat_threat_climate_energytransport)


## Climate + Habitat
#test
sum(data.final$cat_threat_climate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_climate[i]==1 & data.final$cat_threat_habitat[i]==1) {
    data.final$cat_threat_climate_habitat[i]=1}
  if(data.final$cat_threat_climate_habitat[i]==1) {
    data.final$cat_threat_climate[i]=0
    data.final$cat_threat_habitat[i]=0
  }
}

#retest
sum(data.final$cat_threat_climate)
sum(data.final$cat_threat_climate_habitat)


## Climate + Invasive Species
#test
sum(data.final$cat_threat_climate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_climate[i]==1 & data.final$cat_threat_inv_spec[i]==1) {
    data.final$cat_threat_climate_inv_spec[i]=1}
  if(data.final$cat_threat_climate_inv_spec[i]==1) {
    data.final$cat_threat_climate[i]=0
    data.final$cat_threat_inv_spec[i]=0
  }
}

#retest
sum(data.final$cat_threat_climate)
sum(data.final$cat_threat_climate_inv_spec)


## Climate + Overexploitation
#test
sum(data.final$cat_threat_climate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_climate[i]==1 & data.final$cat_threat_overexp[i]==1) {
    data.final$cat_threat_climate_overexp[i]=1}
  if(data.final$cat_threat_climate_overexp[i]==1) {
    data.final$cat_threat_climate[i]=0
    data.final$cat_threat_overexp[i]=0
  }
}

#retest
sum(data.final$cat_threat_climate)
sum(data.final$cat_threat_climate_overexp)


## Climate + Pollution
#test
sum(data.final$cat_threat_climate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_climate[i]==1 & data.final$cat_threat_pollution[i]==1) {
    data.final$cat_threat_climate_pollution[i]=1}
  if(data.final$cat_threat_climate_pollution[i]==1) {
    data.final$cat_threat_climate[i]=0
    data.final$cat_threat_pollution[i]=0
  }
}

#retest
sum(data.final$cat_threat_climate)
sum(data.final$cat_threat_climate_pollution)


## Energy/Transportation + Habitat
#test
sum(data.final$cat_threat_energytransport)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_energytransport[i]==1 & data.final$cat_threat_habitat[i]==1) {
    data.final$cat_threat_energytransport_habitat[i]=1}
  if(data.final$cat_threat_energytransport_habitat[i]==1) {
    data.final$cat_threat_energytransport[i]=0
    data.final$cat_threat_habitat[i]=0
  }
}

#retest
sum(data.final$cat_threat_energytransport)
sum(data.final$cat_threat_energytransport_habitat)


## Energy/Transportation + Invasive Species
#test
sum(data.final$cat_threat_energytransport)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_energytransport[i]==1 & data.final$cat_threat_inv_spec[i]==1) {
    data.final$cat_threat_energytransport_inv_spec[i]=1}
  if(data.final$cat_threat_energytransport_inv_spec[i]==1) {
    data.final$cat_threat_energytransport[i]=0
    data.final$cat_threat_inv_spec[i]=0
  }
}

#retest
sum(data.final$cat_threat_energytransport)
sum(data.final$cat_threat_energytransport_inv_spec)


## Energy/Transportation + Overexploitation
#test
sum(data.final$cat_threat_energytransport)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_energytransport[i]==1 & data.final$cat_threat_overexp[i]==1) {
    data.final$cat_threat_energytransport_overexp[i]=1}
  if(data.final$cat_threat_energytransport_overexp[i]==1) {
    data.final$cat_threat_energytransport[i]=0
    data.final$cat_threat_overexp[i]=0
  }
}

#retest
sum(data.final$cat_threat_energytransport)
sum(data.final$cat_threat_energytransport_overexp)


## Energy/Transportation + Pollution
#test
sum(data.final$cat_threat_energytransport)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_energytransport[i]==1 & data.final$cat_threat_pollution[i]==1) {
    data.final$cat_threat_energytransport_pollution[i]=1}
  if(data.final$cat_threat_energytransport_pollution[i]==1) {
    data.final$cat_threat_energytransport[i]=0
    data.final$cat_threat_pollution[i]=0
  }
}

#retest
sum(data.final$cat_threat_energytransport)
sum(data.final$cat_threat_energytransport_pollution)


## Habitat + Invasive Species
#test
sum(data.final$cat_threat_habitat)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_habitat[i]==1 & data.final$cat_threat_inv_spec[i]==1) {
    data.final$cat_threat_habitat_inv_spec[i]=1}
  if(data.final$cat_threat_habitat_inv_spec[i]==1) {
    data.final$cat_threat_habitat[i]=0
    data.final$cat_threat_inv_spec[i]=0
  }
}

#retest
sum(data.final$cat_threat_habitat)
sum(data.final$cat_threat_habitat_inv_spec)


## Habitat + Overexploitation
#test
sum(data.final$cat_threat_habitat)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_habitat[i]==1 & data.final$cat_threat_overexp[i]==1) {
    data.final$cat_threat_habitat_overexp[i]=1}
  if(data.final$cat_threat_habitat_overexp[i]==1) {
    data.final$cat_threat_habitat[i]=0
    data.final$cat_threat_overexp[i]=0
  }
}

#retest
sum(data.final$cat_threat_habitat)
sum(data.final$cat_threat_habitat_overexp)


## Habitat + Pollution
#test
sum(data.final$cat_threat_habitat)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_habitat[i]==1 & data.final$cat_threat_pollution[i]==1) {
    data.final$cat_threat_habitat_pollution[i]=1}
  if(data.final$cat_threat_habitat_pollution[i]==1) {
    data.final$cat_threat_habitat[i]=0
    data.final$cat_threat_pollution[i]=0
  }
}

#retest
sum(data.final$cat_threat_habitat)
sum(data.final$cat_threat_habitat_pollution)


## Invasive Species + Overexploitation
#test
sum(data.final$cat_threat_inv_spec)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_inv_spec[i]==1 & data.final$cat_threat_overexp[i]==1) {
    data.final$cat_threat_inv_spec_overexp[i]=1}
  if(data.final$cat_threat_inv_spec_overexp[i]==1) {
    data.final$cat_threat_inv_spec[i]=0
    data.final$cat_threat_overexp[i]=0
  }
}

#retest
sum(data.final$cat_threat_inv_spec)
sum(data.final$cat_threat_inv_spec_overexp)


## Invasive Species + Pollution
#test
sum(data.final$cat_threat_inv_spec)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_inv_spec[i]==1 & data.final$cat_threat_pollution[i]==1) {
    data.final$cat_threat_inv_spec_pollution[i]=1}
  if(data.final$cat_threat_inv_spec_pollution[i]==1) {
    data.final$cat_threat_inv_spec[i]=0
    data.final$cat_threat_pollution[i]=0
  }
}

#retest
sum(data.final$cat_threat_inv_spec)
sum(data.final$cat_threat_inv_spec_pollution)


## Overexploitation + Pollution
#test
sum(data.final$cat_threat_overexp)

for(i in 1:nrow(data.final)){
  if(data.final$cat_threat_overexp[i]==1 & data.final$cat_threat_pollution[i]==1) {
    data.final$cat_threat_overexp_pollution[i]=1}
  if(data.final$cat_threat_overexp_pollution[i]==1) {
    data.final$cat_threat_overexp[i]=0
    data.final$cat_threat_pollution[i]=0
  }
}

#retest
sum(data.final$cat_threat_overexp)
sum(data.final$cat_threat_overexp_pollution)
```

### Multi-category Filtering: Taxa Category
```{r}
taxa <- data.final[,c(1:10, 30:36)]

taxa_other <- data.final %>% 
  mutate(sum_taxa = rowSums(.[30:36])) %>%
  filter(sum_taxa > 2)


## Adding 2-taxa categories and other
data.final <- data.final %>%
  mutate(cat_taxa_amphibian_bird = 0,
         cat_taxa_amphibian_fish = 0,
         cat_taxa_amphibian_invertebrate = 0,
         cat_taxa_amphibian_mammal = 0,
         cat_taxa_amphibian_plant = 0,
         cat_taxa_amphibian_reptile = 0,
         cat_taxa_bird_fish = 0,
         cat_taxa_bird_invertebrate = 0,
         cat_taxa_bird_mammal = 0,
         cat_taxa_bird_plant = 0,
         cat_taxa_bird_reptile = 0,
         cat_taxa_fish_invertebrate = 0,
         cat_taxa_fish_mammal = 0,
         cat_taxa_fish_plant = 0,
         cat_taxa_fish_reptile = 0,
         cat_taxa_invertebrate_mammal = 0,
         cat_taxa_invertebrate_plant = 0,
         cat_taxa_invertebrate_reptile = 0,
         cat_taxa_mammal_plant = 0,
         cat_taxa_mammal_reptile = 0,
         cat_taxa_plant_reptile = 0,
         cat_taxa_other = 0)

#### Filling Multiple Taxa cells ####

## Other Taxa (>2 classifications)
#test
sum(data.final$cat_taxa_plant)

for(i in 1:nrow(data.final)) {
  if(sum(data.final[i,30:36]) > 2) {
    data.final$cat_taxa_other[i] = 1
    data.final[i,30:36] = 0
  }
}

#retest
sum(data.final$cat_taxa_plant)
sum(data.final$cat_taxa_other) # matches nrow of taxa_other DF


## amphibian + bird
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_bird[i]==1) {
    data.final$cat_taxa_amphibian_bird[i]=1}
  if(data.final$cat_taxa_amphibian_bird[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_bird[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_bird)


## amphibian + fish
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_fish[i]==1) {
    data.final$cat_taxa_amphibian_fish[i]=1}
  if(data.final$cat_taxa_amphibian_fish[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_fish[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_fish)


## amphibian + invertebrate
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_invertebrate[i]==1) {
    data.final$cat_taxa_amphibian_invertebrate[i]=1}
  if(data.final$cat_taxa_amphibian_invertebrate[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_invertebrate[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_invertebrate)


## amphibian + mammal
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_mammal[i]==1) {
    data.final$cat_taxa_amphibian_mammal[i]=1}
  if(data.final$cat_taxa_amphibian_mammal[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_mammal[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_mammal)


## amphibian + plant
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_plant[i]==1) {
    data.final$cat_taxa_amphibian_plant[i]=1}
  if(data.final$cat_taxa_amphibian_plant[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_plant[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_plant)


## amphibian + reptile
#test
sum(data.final$cat_taxa_amphibian)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_amphibian[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_amphibian_reptile[i]=1}
  if(data.final$cat_taxa_amphibian_reptile[i]==1) {
    data.final$cat_taxa_amphibian[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_amphibian)
sum(data.final$cat_taxa_amphibian_reptile)


## bird + fish
#test
sum(data.final$cat_taxa_bird)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_bird[i]==1 & data.final$cat_taxa_fish[i]==1) {
    data.final$cat_taxa_bird_fish[i]=1}
  if(data.final$cat_taxa_bird_fish[i]==1) {
    data.final$cat_taxa_bird[i]=0
    data.final$cat_taxa_fish[i]=0
  }
}

#retest
sum(data.final$cat_taxa_bird)
sum(data.final$cat_taxa_bird_fish)


## bird + invertebrate
#test
sum(data.final$cat_taxa_bird)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_bird[i]==1 & data.final$cat_taxa_invertebrate[i]==1) {
    data.final$cat_taxa_bird_invertebrate[i]=1}
  if(data.final$cat_taxa_bird_invertebrate[i]==1) {
    data.final$cat_taxa_bird[i]=0
    data.final$cat_taxa_invertebrate[i]=0
  }
}

#retest
sum(data.final$cat_taxa_bird)
sum(data.final$cat_taxa_bird_invertebrate)


## bird + mammal
#test
sum(data.final$cat_taxa_bird)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_bird[i]==1 & data.final$cat_taxa_mammal[i]==1) {
    data.final$cat_taxa_bird_mammal[i]=1}
  if(data.final$cat_taxa_bird_mammal[i]==1) {
    data.final$cat_taxa_bird[i]=0
    data.final$cat_taxa_mammal[i]=0
  }
}

#retest
sum(data.final$cat_taxa_bird)
sum(data.final$cat_taxa_bird_mammal)


## bird + plant
#test
sum(data.final$cat_taxa_bird)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_bird[i]==1 & data.final$cat_taxa_plant[i]==1) {
    data.final$cat_taxa_bird_plant[i]=1}
  if(data.final$cat_taxa_bird_plant[i]==1) {
    data.final$cat_taxa_bird[i]=0
    data.final$cat_taxa_plant[i]=0
  }
}

#retest
sum(data.final$cat_taxa_bird)
sum(data.final$cat_taxa_bird_plant)


## bird + reptile
#test
sum(data.final$cat_taxa_bird)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_bird[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_bird_reptile[i]=1}
  if(data.final$cat_taxa_bird_reptile[i]==1) {
    data.final$cat_taxa_bird[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_bird)
sum(data.final$cat_taxa_bird_reptile)


## fish + invertebrate
#test
sum(data.final$cat_taxa_fish)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_fish[i]==1 & data.final$cat_taxa_invertebrate[i]==1) {
    data.final$cat_taxa_fish_invertebrate[i]=1}
  if(data.final$cat_taxa_fish_invertebrate[i]==1) {
    data.final$cat_taxa_fish[i]=0
    data.final$cat_taxa_invertebrate[i]=0
  }
}

#retest
sum(data.final$cat_taxa_fish)
sum(data.final$cat_taxa_fish_invertebrate)


## fish + mammal
#test
sum(data.final$cat_taxa_fish)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_fish[i]==1 & data.final$cat_taxa_mammal[i]==1) {
    data.final$cat_taxa_fish_mammal[i]=1}
  if(data.final$cat_taxa_fish_mammal[i]==1) {
    data.final$cat_taxa_fish[i]=0
    data.final$cat_taxa_mammal[i]=0
  }
}

#retest
sum(data.final$cat_taxa_fish)
sum(data.final$cat_taxa_fish_mammal)


## fish + plant
#test
sum(data.final$cat_taxa_fish)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_fish[i]==1 & data.final$cat_taxa_plant[i]==1) {
    data.final$cat_taxa_fish_plant[i]=1}
  if(data.final$cat_taxa_fish_plant[i]==1) {
    data.final$cat_taxa_fish[i]=0
    data.final$cat_taxa_plant[i]=0
  }
}

#retest
sum(data.final$cat_taxa_fish)
sum(data.final$cat_taxa_fish_plant)


## fish + reptile
#test
sum(data.final$cat_taxa_fish)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_fish[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_fish_reptile[i]=1}
  if(data.final$cat_taxa_fish_reptile[i]==1) {
    data.final$cat_taxa_fish[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_fish)
sum(data.final$cat_taxa_fish_reptile)


## invertebrate + mammal
#test
sum(data.final$cat_taxa_invertebrate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_invertebrate[i]==1 & data.final$cat_taxa_mammal[i]==1) {
    data.final$cat_taxa_invertebrate_mammal[i]=1}
  if(data.final$cat_taxa_invertebrate_mammal[i]==1) {
    data.final$cat_taxa_invertebrate[i]=0
    data.final$cat_taxa_mammal[i]=0
  }
}

#retest
sum(data.final$cat_taxa_invertebrate)
sum(data.final$cat_taxa_invertebrate_mammal)


## invertebrate + plant
#test
sum(data.final$cat_taxa_invertebrate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_invertebrate[i]==1 & data.final$cat_taxa_plant[i]==1) {
    data.final$cat_taxa_invertebrate_plant[i]=1}
  if(data.final$cat_taxa_invertebrate_plant[i]==1) {
    data.final$cat_taxa_invertebrate[i]=0
    data.final$cat_taxa_plant[i]=0
  }
}

#retest
sum(data.final$cat_taxa_invertebrate)
sum(data.final$cat_taxa_invertebrate_plant)


## invertebrate + reptile
#test
sum(data.final$cat_taxa_invertebrate)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_invertebrate[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_invertebrate_reptile[i]=1}
  if(data.final$cat_taxa_invertebrate_reptile[i]==1) {
    data.final$cat_taxa_invertebrate[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_invertebrate)
sum(data.final$cat_taxa_invertebrate_reptile)


## mammal + plant
#test
sum(data.final$cat_taxa_mammal)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_mammal[i]==1 & data.final$cat_taxa_plant[i]==1) {
    data.final$cat_taxa_mammal_plant[i]=1}
  if(data.final$cat_taxa_mammal_plant[i]==1) {
    data.final$cat_taxa_mammal[i]=0
    data.final$cat_taxa_plant[i]=0
  }
}

#retest
sum(data.final$cat_taxa_mammal)
sum(data.final$cat_taxa_mammal_plant)


## mammal + reptile
#test
sum(data.final$cat_taxa_mammal)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_mammal[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_mammal_reptile[i]=1}
  if(data.final$cat_taxa_mammal_reptile[i]==1) {
    data.final$cat_taxa_mammal[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_mammal)
sum(data.final$cat_taxa_mammal_reptile)


## plant + reptile
#test
sum(data.final$cat_taxa_plant)

for(i in 1:nrow(data.final)){
  if(data.final$cat_taxa_plant[i]==1 & data.final$cat_taxa_reptile[i]==1) {
    data.final$cat_taxa_plant_reptile[i]=1}
  if(data.final$cat_taxa_plant_reptile[i]==1) {
    data.final$cat_taxa_plant[i]=0
    data.final$cat_taxa_reptile[i]=0
  }
}

#retest
sum(data.final$cat_taxa_plant)
sum(data.final$cat_taxa_plant_reptile)
```

### Export Multi-Category Data for Analysis and Visualization
```{r, eval=FALSE}
write_csv(data.final, "~/Desktop/master_output_data_analysis_v15_filtered.csv")
```
