# Data Analysis and Visualization {dataVis}

```{r include=FALSE}
library(tidyverse)
library(scales)
library(ggsci)
```
```{r, echo = FALSE}
# Data Import/Prep
data <- read_csv("~/Desktop/master_output_data_analysis_v15_filtered.csv") %>%
   as.data.frame() %>%
   dplyr::select(c(1:42, 49:109, 43:48))

data.multi <- read_csv("~/Desktop/master_output_data_analysis_v15_mutlicategories.csv") %>%
   as.data.frame()

maxwell <- read_csv("~/Desktop/Conservation Lab//Maxwell et al table S2.csv") %>%
  as.data.frame()
```
```{r, echo = FALSE}
## Category Lists
clim.energy_cat <- names(data)[c(37,38,66:74,81)]         # climate/energy/combos/other

biodiv.bio.taxa.hab_cat <- names(data)[c(19:28,43, 30:36, 82:103)]
biodiv.bio.taxa.hab_cat_multi <- names(data.multi)[c(19:28,30:36)]

u.threat_cat <- names(data)[37:42]
u.taxa_cat <- names(data)[30:36]

all.threat_cat <- append(u.threat_cat, names(data)[66:81])
all.taxa_cat <- append(u.taxa_cat, names(data)[82:103])

u.threat_list <- c("Climate Change", 
                   "Energy & Transportation",
                   "Habitat Loss",
                   "Invasive Species",
                   "Overexploitation",
                   "Pollution")
u.taxa_list <- c("Amphibians",
                 "Birds",
                 "Fish",
                 "Invertebrates",
                 "Mammals",
                 "Plants",
                 "Reptiles")
```
```{r, echo = FALSE}
## Colors
cat_colors22 <- c("#990000", "#BF1E1E", "#993399", "#4C0099", "#000099",
                  "#2E10C3", "#3685B3", "#1AB3B3", "#8FC04F", "#09A256", 
                  "#006400", "#FFF928", "#C6C600", "#FFA200", "#CE8300", 
                  "#D0471E", "#9F8109", "#835506", "#317575", "#808080",
                  "black", "white")
barplot(1:22, col = cat_colors22, names.arg = 1:22)

cat_colors7 <- cat_colors22[c(1, 5, 20, 7, 18, 11, 19)]
barplot(1:7, col = cat_colors7)
```

## Exploratory Visuals
### Grants with No Keywords
This is a plot featuring the 50 largest grants that have no keywords associated with them.
```{r noKeywordPlot, fig.cap='Large grants with no keywords', fig.asp=.75, fig.align='center'}
no_keywords <- data %>% 
  filter(Keywords =="['-']") %>%
  arrange(desc(Grant.Amount))
no_keywords_plot <- ggplot(no_keywords[1:50,],
                            aes(x = factor(row.names(no_keywords[1:50,]), levels = 1:50),
                                y = Grant.Amount)) + 
  geom_col() +
  theme_minimal() +
  labs(title = "No Keyword Biggest Grants (50)",
       x = "Rank",
       y = "Money ($US)") +
  scale_y_continuous(labels = scales::comma)
no_keywords_plot
```

### All Grants with Funding  

Here we lengthen the data to more properly examine and visualize where the money from each grant is being allocated.
```{r}
data <- filter(data, Keywords != "['-']") %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002)
data.multi <- filter(data.multi, Keywords != "['-']") %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002)

money.data <- data[,11:(length(data)-6)]*data[,7]
money.data <- mutate(money.data, Year.Authorized = data$Year.Authorized)

by.year.money <- money.data %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_taxa_other), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))
```

### Bio Subset
Here we remove those climate/energy grants that do not contain biodiversity conservation keywords.
```{r}
# Single Category Data
data_bio.1 <- data %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat), any_vars(.> 0)) 
data_non.bio.1 <- data %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat), all_vars(. == 0))

data_bio.2 <- data_non.bio.1 %>%
  filter_at(vars(clim.energy_cat), all_vars(.==0))
data_non.bio.2 <- data_non.bio.1 %>%
  filter_at(vars(clim.energy_cat), any_vars(.>0))

data_bio <- bind_rows(data_bio.1, data_bio.2)
data_non.bio <- data_non.bio.2

# Multi-category Data
data.multi_bio.1 <- data.multi %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat_multi), any_vars(.> 0)) 
data.multi_non.bio.1 <- data.multi %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat_multi), all_vars(. == 0))

data.multi_bio.2 <- data.multi_non.bio.1 %>%
  filter_at(vars(cat_threat_climate, cat_threat_energytransport),
            all_vars(.==0))
data.multi_non.bio.2 <- data.multi_non.bio.1 %>%
  filter_at(vars(cat_threat_climate, cat_threat_energytransport),
            any_vars(.>0))

data.multi_bio <- bind_rows(data.multi_bio.1, data.multi_bio.2)
data.multi_non.bio <- data.multi_non.bio.2
```

### Money Data
Here we further subset the lengthened dataset featuring all grants that have associated funding and break them down at an annual resolution Those subsets include the amount of money allocated to each category per year, the amount of non-biodiversity funding by year, and the amount of money across multi-category grants per year.
```{r}
# Single Category Data
money.data_bio <- data_bio[,11:(length(data_bio)-6)]*data_bio[,7]
money.data_bio <- mutate(money.data_bio,
                         Year.Authorized = data_bio$Year.Authorized,
                         Num.Thr = data_bio$Num.Thr)

by.year.money_bio <- money.data_bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_taxa_other), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Year.Authorized) %>%
  mutate(ann.percent = Amount/sum(Amount)) %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))

money.data_non.bio <- data_non.bio[,11:(length(data_non.bio)-6)]*data_non.bio[,7]
money.data_non.bio <- mutate(money.data_non.bio, 
                             Year.Authorized = data_non.bio$Year.Authorized,
                             Num.Thr = data_non.bio$Num.Thr)

by.year.money_non.bio <- money.data_non.bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_threat_overexp_pollution), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))

# Multi-category Data
money.data.multi_bio <- data.multi_bio[,11:(length(data.multi_bio)-6)]*data.multi_bio[,7]
money.data.multi_bio <- mutate(money.data.multi_bio, 
                               Year.Authorized = data.multi_bio$Year.Authorized,
                               Num.Thr = data.multi_bio$Num.Thr)

by.year.money.multi_bio <- money.data.multi_bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_threat_pollution), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))
```

## Manuscript Figures
### Figure 1: Proportion of Money vs. Species Threatened
**INSERT FIGURE CAPTION FROM MANUSCRIPT**
```{r fig-1, fig.cap='Figure 1: Proportion of Money vs. Species Threatened', fig.asp=.75, fig.align='center'}
fig.1.money_bio <- by.year.money_bio %>%
  filter(Category %in% u.threat_cat) %>%
  group_by(Category) %>%
  summarise(Total = sum(Amount))

fig.1.species <- maxwell %>% 
  dplyr::select(Category, n_species_threatened) %>%
  group_by(Category) %>%
  summarise(Total = sum(n_species_threatened)) 

rawFrame <- left_join(fig.1.money_bio, fig.1.species, by = "Category") %>%
  mutate(rawMoney = Total.x) %>%   # Total Money
  mutate(rawSpecies = Total.y) %>% # Total Species
  dplyr::select(Category, rawMoney, rawSpecies)

fig.1.data_bio <- left_join(fig.1.money_bio, fig.1.species, by = "Category") %>%
  mutate(Money = (Total.x/sum(Total.x))*100) %>%   # Percentage of Total Money
  mutate(Species = (Total.y/sum(Total.y))*100) %>% # Percentage of Total Species
  dplyr::select(Category, Money, Species) %>%
  pivot_longer(c(Money, Species), names_to = "is.Money", values_to = "Amount") %>%
  left_join(rawFrame, by = "Category") %>%
  mutate(rawMoney = ifelse(is.Money == "Species", NA, rawMoney)) %>%
  mutate(rawSpecies= ifelse(is.Money == "Money", NA, rawSpecies)) %>%
  replace_na(list(rawMoney = "", rawSpecies = "")) %>%
  unite("rawValue", rawMoney:rawSpecies, sep = "", remove = FALSE) %>%
  mutate(rawValue = as.numeric(rawValue),
         rawLabels = comma(rawValue))
fig.1.data_bio$rawLabels[c(seq(1,11,2))] <- str_glue("${round(rawFrame$rawMoney/1e6)}m")
# fig.1.data_bio$rawLabels[c(seq(2,12,2))] <- rawFrame$rawSpecies

fig.1.plot_bio <- ggplot(fig.1.data_bio, 
                      aes(fill=is.Money, 
                          y=Amount, 
                          x=Category)) + 
  geom_col(position="dodge") +
  labs(x="Threat Category", 
       y="Percentage") + 
       # title="% Money Allocated vs. % Species Threatened (IUCN) by Unique Threats") +
  scale_x_discrete(labels=u.threat_list) +
  scale_fill_manual(# values = c("#556B2F", "#8f3c36"),
                    values = alpha(c("darkgreen", "orangered4"), 0.9),
                    # values = c("#556B2F", "#708090"),
                    labels = c("Money ($USD Million)", "# Threatened Species")) +
  geom_text(position = position_dodge(width= 1), 
            aes(label = rawLabels, hjust = 0.5, vjust = -0.5), angle=0, size=6) +
  ylim(0, 50) +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = "bold"), 
        axis.ticks.length.x = unit(.5, "cm"),
        axis.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.key.size = unit(1,"cm"),
        legend.text = element_text(size = 12))

# ggsave(fig.1.plot_bio, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/money_species_comparison.png", width=15, height=7,dpi = 300)

fig.1.plot_bio
```

### Figure 2: Amount of Money Allocated Across Unique Threat Categories per Year
**INSERT FIGURE CAPTION FROM MANUSCRIPT**
```{r fig-2, fig.cap='Figure 2: Amount of Money Allocated Across Unique Threat Categories per Year', fig.asp=.75, fig.align='center'}
threat_temporal.data <- by.year.money_bio %>%
  filter(Category %in% u.threat_cat)

threat_temporal.plot <- ggplot(threat_temporal.data, 
                aes(y = Amount, 
                    x = Year.Authorized,
                    color = Category)) +
  #geom_point() +
  geom_line(size = 2, linetype = 1) +
  # geom_smooth(size = 1, aes(fill = Category), show.legend = FALSE) +
  theme_minimal() +
  scale_color_manual(values = cat_colors22[c(1,13,11,20,5,18)],
                     name="Threat Category",
                     labels=u.threat_list) +
  scale_fill_manual(values = cat_colors7) +
  labs(x="Year", 
       y="Money ($US Million)", 
       title="Amount of Money Allocated Across Unique Threat Categories per Year") +
  scale_y_continuous(#labels = waiver(),
                     breaks = c(0,50e6,10e7),
                     labels = c("0","50","100"),
                     position = "right") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, face = "bold"),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.key.width = unit(1,"cm"),
        # legend.box.spacing = unit(1,"cm"),
        legend.spacing.x = unit(1,"cm"),
        legend.position = "bottom")
        # legend.background = element_rect(color = "black", size = 1.2, fill = NA))

# ggsave(threat_temporal.plot, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/threat_temporal.png", width=12, height=7,dpi = 300)


threat_temporal.plot
```

### Figure 3: Taxa/Threat Maxwell Comparison
**INSERT FIGURE CAPTION FROM MANUSCRIPT**
#### Coercing Maxwell's species designations to merge with ours
We are merging the various taxa catgories used by Maxwell et al. to examine the amount of funding disbursed to various taxa. Maxwell et al.'s taxonic designations are at a higher resolution than the what we need, so we are altering them to fit the taxonomic scale we're operating at.
```{r}
maxwell.s4 <- read_csv("~/Columbia E3B/Conservation Group/Maxwell et al table S4.csv")
names(maxwell.s4)

# Our species: Amphibians, Birds, Fish, Invertebrates, Mammals, Plants, Reptiles

our.species <- maxwell.s4 %>%
  slice(1:10) %>%
  mutate(Amphibians = Amphibia,
         Birds = Aves,
         Fish = Actinopterygii + Chondrichthyes,
         Invertebrates = Anthozoa + Gastropoda + Hydrozoa + Malacostraca + Myxini,
         Mammals = Mammalia,
         Plants = Cycadopsida + Liliopsida + Magnoliopsida + Pinopsida,
         Reptiles = Reptilia) %>%
  dplyr::select(c(2,3, 19:ncol(.), 18))
our.species
```

#### Figure 3
```{r}
# Preparing maxwell data
fig.3.maxwell_data <- our.species %>% 
  dplyr::select(2:11) %>%
  group_by(Category) %>%
  summarise_at(vars(Total:Reptiles), sum) %>%
  dplyr::select(3:length(.)) %>%
  mutate(Total.maxwell = rowSums(.)) %>%
  dplyr::select(8,1:7) %>%
  transmute_at(vars(-Total.maxwell), ~ (./Total.maxwell)*100) %>%
  mutate(Threat = u.threat_list) %>%
  dplyr::select(8,1:7) %>%
  pivot_longer(Amphibians:Reptiles, names_to = "Taxa", values_to = "Percent") %>%
  mutate(Source = "IUCN", direction = -1, alpha = 0.9)


# Preparing money data
taxa_climate <- money.data_bio %>%
  filter(cat_threat_climate > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_climate") %>%
  `colnames<-` (u.taxa_list)

taxa_energy <- money.data_bio %>%
  filter(cat_threat_energytransport > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_energytransport") %>%
  `colnames<-` (u.taxa_list)

taxa_habitat <- money.data_bio %>%
  filter(cat_threat_habitat > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_habitat") %>%
  `colnames<-` (u.taxa_list)

taxa_inv <- money.data_bio %>%
  filter(cat_threat_inv_spec > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_inv_spec") %>%
  `colnames<-` (u.taxa_list)

taxa_overexp <- money.data_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_overexp") %>%
  `colnames<-` (u.taxa_list)

taxa_pollution <- money.data_bio %>%
  filter(cat_threat_pollution > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_pollution") %>%
  `colnames<-` (u.taxa_list)

fig.3.money_data <- bind_rows(taxa_climate, taxa_energy, taxa_habitat, 
                         taxa_inv, taxa_overexp, taxa_pollution) %>%
  transmute_all(~.*100) %>%
  mutate(Threat = u.threat_list) %>%
  dplyr::select(8, 1:7) %>%
  pivot_longer(Amphibians:Reptiles, names_to = "Taxa", values_to = "Percent") %>%
  mutate(Source = "Money", direction = 1, alpha = 1)

### 1 plot
fig.3.data <- bind_rows(fig.3.money_data, fig.3.maxwell_data) %>%
  mutate(Percent = Percent * direction,
         Taxa = factor(Taxa, 
                       levels = c("Reptiles", "Amphibians", "Invertebrates", 
                                  "Birds", "Mammals", "Fish", "Plants"),
                       ordered = TRUE),
         Threat = factor(Threat,
                         levels = c("Climate Change", "Overexploitation",
                                    "Habitat Loss", "Energy & Transportation",
                                    "Pollution", "Invasive Species"),
                         ordered = TRUE)
  )

fig.3 <- ggplot(fig.3.data %>% filter(Taxa != "Reptiles"),
                aes(x = Taxa, y = Percent, fill = Taxa)) + 
  geom_col(show.legend = F , color = "black") +
  geom_hline(yintercept = 0, size = 1.2) +
  scale_fill_manual(values = cat_colors22[c(9,20,19,15,7,11)],
                    name = "Taxa",
                    labels = u.taxa_list) +
  facet_wrap(~Threat) +
  theme_bw() +
  theme(axis.ticks.y= element_blank(),
        # axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_blank(),
        # axis.text.y = element_text(size = 12),
        # axis.title = element_text(size = 16, face = "bold"),
        axis.ticks.x=element_line(color=NA),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(size = 12, color = "white", face = "bold")) +
  ylim(-90,90) +
  ggtitle("Percent of Taxa Threatened (IUCN) by vs. Funded for Unique Threats") +
  ylab("% Taxa Threatened   |    % Funding to Taxa") +
  coord_flip()

# ggsave(fig.3, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/taxa_threat_comparison.png", width=12, height=7,dpi = 300)

fig.3
```

### Figure 4: Network Analysis
# Figure 4: Threats Network Analysis
The code presented here for Figure 4 generates the pie charts for each threat category that can be seen in the full network analysis figure. The actual network analysis was conducted separately, in **WHERE WAS THIS CONDUCTED?**.

## Threat Category Overlap
```{r, include=FALSE}
threat.net1 <- data.multi_bio %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002) %>%
  dplyr::select(cat_threat_climate:cat_threat_pollution)
threat.net1 <- map2_df(threat.net1, colnames(threat.net1), ~  c(NA, .y)[.x +1]) # Change the binary 1 values into the column heaers
threat.net2 <- data.multi_bio %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002) %>%
  dplyr::select(Grant.Amount, Num.Thr)
threat.net3 <- bind_cols(threat.net2, threat.net1) # Join the column-header data with the grant amount/number of threats per grant

threat.net3 <- threat.net3 %>%
  filter(Num.Thr != 0 & Num.Thr <= 2) %>%
  unite("z", sep = ",", cat_threat_climate:cat_threat_pollution, na.rm = TRUE, remove = FALSE) %>% # Bring together all columns into a single column, separate contents by a comma
  dplyr::select(Grant.Amount, Num.Thr, z) %>%
  separate(z, sep = ",", c("A", "B")) %>% # Break united column into two
  mutate(B = if_else(is.na(B), A, B)) %>% # Have all NA's = Column A
  mutate(A = case_when(
      A == "cat_threat_climate" ~ "Climate Change",
      A == "cat_threat_energytransport" ~ "Energy & Transportation",
      A == "cat_threat_habitat" ~ "Habitat Loss",
      A == "cat_threat_inv_spec" ~ "Invasive Species",
      A == "cat_threat_overexp" ~ "Overexploitation",
      A == "cat_threat_pollution" ~ "Pollution")) %>%
  mutate(B = case_when(
      B == "cat_threat_climate" ~ "Climate Change",
      B == "cat_threat_energytransport" ~ "Energy & Transportation",
      B == "cat_threat_habitat" ~ "Habitat Loss",
      B == "cat_threat_inv_spec" ~ "Invasive Species",
      B == "cat_threat_overexp" ~ "Overexploitation",
      B == "cat_threat_pollution" ~ "Pollution")) %>%
  group_by(A, B) %>% # group by the combinations between column A and column B, which contain one or two threats
  summarize(n = sum(Grant.Amount)) %>%
  filter(A != B) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

# write_csv(threat.net3, "~/Columbia E3B/Conservation Group/Output V15.2/Tables/threat_overlap_bio.csv")

threat.net.fig <- threat.net3 %>%
  ggplot(aes(x = A, y = B)) +
  geom_tile(aes(fill = log10(n)), color = "black") +
  geom_text(aes(label = n, fontface = "bold"), color = "white") +
  scale_fill_gradient(low = "lightgray", high = "black") +
  theme_bw() +
  labs(x="",
       y = "",
       title = "Amount of Money ($USD) in Single and Paired Threat Categories") +
  theme(legend.position = "none",
        panel.grid.major = element_blank())

threat.net.fig
```

## Pie Charts
```{r}
blank_theme <- theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        plot.title=element_text(size=14, face="bold"))
```

### # of Threats within Threat categories (number of grants, not $)
#### Climate
```{r}
pie.Nthreats_climate <- money.data.multi_bio %>%
  filter(cat_threat_climate > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount),
         alpha = seq(0.2,1,0.2))

plot.pie.Nthreats_climate <- ggplot(pie.Nthreats_climate,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[1], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_manual(values = c("#a50f15","#de2d26","#fb6a4a","#fcae91","#fee5d9")) +
  # scale_fill_viridis_d(name = "Number of Threats", option = "D") +
  labs(title="Proportion of Climate Change Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

# ggsave(plot.pie.Nthreats_climate, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_climate.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_climate
```


#### Energy/Transportation
```{r}
pie.Nthreats_energytransport <- money.data.multi_bio %>%
  filter(cat_threat_energytransport > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_energytransport <- ggplot(pie.Nthreats_energytransport,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "black") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[13], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Energy/Transportation Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

# ggsave(plot.pie.Nthreats_energytransport, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_energytransport.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_energytransport
```


#### Habitat Loss
```{r}
pie.Nthreats_habitat <- money.data.multi_bio %>%
  filter(cat_threat_habitat > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_habitat <- ggplot(pie.Nthreats_habitat,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
    scale_fill_manual(values = seq_gradient_pal(cat_colors22[11], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats", end = 0.8) +
  labs(title="Proportion of Habitat Loss Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4)),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_habitat, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_habitat.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_habitat
```


#### Invasive Species
```{r}
pie.Nthreats_inv_spec <- money.data.multi_bio %>%
  filter(cat_threat_inv_spec > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_inv_spec <- ggplot(pie.Nthreats_inv_spec,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "black") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[20], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Invasive Species Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_inv_spec, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_inv_spec.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_inv_spec
```

#### Overexploitation
```{r}
pie.Nthreats_overexp <- money.data.multi_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_overexp <- ggplot(pie.Nthreats_overexp,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[5], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Overexploitation Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                # label = paste0(round(perc*100,1), "%")),
                label = str_glue("{round(perc*100,1)}% ({totalCount})")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_overexp, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_overexp.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_overexp
```

##### Prop Overxploitation funding
```{r}
pie.threats_overexp <- money.data.multi_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(Num.Thr, cat_threat_overexp) %>%
  group_by(Num.Thr) %>%
  summarize(total = sum(cat_threat_overexp)) %>%
  mutate(perc = total/sum(total))
```


#### Pollution
```{r}
pie.Nthreats_pollution <- money.data.multi_bio %>%
  filter(cat_threat_pollution > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_pollution <- ggplot(pie.Nthreats_pollution,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[18], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Pollution Funding Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_pollution, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_pollution.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_pollution
```

## Supplemental Figures
### Figure S1: Total Biodiversity/Climate Change Grantmaking by Top 1000 Grantmakers
```{r fig-s1, fig.cap='Figure S1: Grant Funding by Top 1,000 Grantmakers', fig.asp=.75, fig.align='center'}
donor_Rank <- c(" ", "Top 50", "Top 100", "Top 200", "Top 300", "Top 400", "Top 500", "Top 600", "Top 700", "Top 800", "Top 900", "Top 1000")
sum <- c(0, 14396513202, 16440571398, 18171144637, 18975986509, 19490663536, 19851261916, 20128241008, 20348022885, 20525790876, 20670921174, 20793224389)
sum_mildiv <- sum/1000000
axis_Order <- c(1:12)

sum.data <- data.frame(donor_Rank, sum, sum_mildiv, axis_Order)

sum_graph <- ggplot(data = sum.data,
                    mapping = aes(x = axis_Order,
                                  y =  sum_mildiv,
                                  group = 1)
) +
  geom_line() +
  geom_point() +
  geom_area(aes(y=ifelse(sum_mildiv>=0 & sum_mildiv< 19851.26, sum_mildiv, 19851.26)), fill="#696969", alpha = .4) +
  ylim(0, 21500) +
  theme_minimal() +
  xlab("Grantmaker Group") +
  ylab("Sum ($1,000,000USD)") + 
  scale_x_discrete(limits=c(" ", "Top 50", "Top 100", "Top 200", "Top 300", "Top 400", "Top 500", "Top 600", 
                                      "Top 700", "Top 800", "Top 900", "Top 1000"))

sum_graph
```
