---
editor_options: 
  chunk_output_type: console
---

#### Post-processing of grants following validation

Here we remove and filter those grants that may have been erroneously classified, following the process of validation
```{r setup, include=FALSE}
# Load libraries
library(knitr)
library(tidyverse)
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
dat <- read_csv("data\\classified-grants-for-analysis.csv")
```

Here we hard-coded several specific fixes to the classification that repeatedly cropped up during our validation process. All have been checked to have 0 mis-classifications.
```{r}
# Removing grants containing "dinner"
dat <- dat %>%
  filter(str_detect(Description, "dinner") == FALSE)

# Classifying "dissemination" as act_edu
sum(dat$cat_act_edu)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "dissemination") == TRUE) {
    dat$cat_act_edu[i] = 1
  }
}
sum(dat$cat_act_edu)

# Removing threat_energytransport classification for "palm oil"
sum(dat$cat_threat_energytransport)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "palm oil|oil-palm|oil palm") == TRUE) {
    dat$cat_threat_energytransport[i] = 0
  }
}
sum(dat$cat_threat_energytransport)

# Removing hab_other classification for "aquatic" (will be freshwater/marine)
sum(dat$cat_habitat_other)
for(i in 1:nrow(dat)) {
  if(str_detect(dat$Description[i], "aquatic") == TRUE) {
    dat$cat_habitat_other[i] = 0
  }
}
sum(dat$cat_habitat_other)

```

Those grants with no funding amount recorded contributed nothing to this analysis, and were removed.
```{r}
dat.1 <- dat %>% 
  drop_na(Grant.Amount) %>%
  dplyr::select(1:42) 
names(dat.1)
```

We ensured that our study focused on in situ conservation grants by eliminating grants that were directed toward captive management (aquarium/zoo category)

```{r}
for(i in 1:nrow(dat.1)){
  if(dat.1$cat_aquarium_zoo[i]==1){
    dat.1[i,c(11:17, 19:28, 30:42)]=0
  }
}

write_csv(dat,"data\\post-processed-grants-for-analysis.csv")
```
